---
title: "Statistical Inference"
output:
  pdf_document: default
  html_document: default
---

<html>
<body>
  <h1>Statistical Inference on Coursera</h1>
  <table><tr>
    <th>Title:</th>
    <td>Project report on simulation exercise and basic inferential data analysis.</td>
    </tr><tr>
    <th>Author:</th> 
    <td>Chloe Ng</td>
    </tr><tr>
    <th>Overview:</th>
    <td>This report consists of two parts - 1. Simulation of an exponential distribution and its asymtotic approximation 2. Inferential data analysis of ToothGrowth data in R dataset packages.</td>
  </tr></table>
    
  <p>Part 1: <strong>Exponential distribution simulation</strong></p>
  
  <p>The code below plots the distribution of the averages of 40 random variables following exponential distribution with lambda = 0.2, simulated 1000 times.</p>
  
```{r}
if(!require(ggplot2)) install.packages("ggplot2",repos = "http://cran.us.r-project.org")
library("ggplot2")
lambda <- 0.2
n <- 40
no.sim <- 1000
# a quick plot to see how 1000 simulations of 40 exponential r.v. looks like
mainPlot <- qplot(replicate(1000, mean(rexp(40, lambda))),
                  binwidth=0.1, 
                  fill='salmon', alpha = 0.5,
                  show.legend=FALSE,
                  main = 'Distribution of 1000 simulation of means \nof 40 r.v. under exponential distribution') 
mainPlot
```
  <p>Sample Mean vs Theoretical Mean:</p>

```{r}
# create matrix of simulations 
expoMatrix <- matrix(replicate(no.sim,rexp(n, lambda)), nrow = 1000, ncol = 40, byrow = TRUE) # 1000 simulation of 'expo'
# create dataframe for use in ggplot
expo.df <- data.frame(allMeans=apply(expoMatrix, 1,mean))
mean <- apply(expo.df, 2, mean)
# create a frequency histogram
ggplot(data=expo.df, aes(x=allMeans)) + 
    geom_histogram(binwidth =.1, fill='salmon', alpha=0.5) + 
    geom_vline(xintercept = mean, color='darkred',linetype='dashed') + 
    ggtitle('Distribution of 1000 simulation of means \nof 40 r.v. under exponential distribution') + 
    theme(plot.title = element_text(size = 10)) + 
    theme_minimal()
# the mean of simulation is close to the theoretical mean of exponentials
```
  <p>From the above, we have the mean from simulation is <!--rinline mean -->, which is close to 
     the theoretical mean of 5.</p>

  <p>Sample Variance vs Theoretical Variance:</p>
  
```{r}
# create dataframe for ggplot from the existing matrix 
var.df <- data.frame(var=apply(expoMatrix, 1, var))
# calculate mean of all variances 
meanVar <- apply(var.df, 2, mean)
# create histogram and dashed line to indicate mean of variances
ggplot(data=var.df, aes(x=var)) + 
    geom_histogram(binwidth =.5, fill='salmon', alpha=0.5) + 
    geom_vline(xintercept = meanVar, color='darkred',linetype='dashed') + 
    ggtitle('Distribution of 1000 variances of the 40 r.v. under exponential distribution') + 
    theme(plot.title = element_text(size = 10)) + 
    theme_minimal()
```
    
  <p>From the above, we have the mean of variances from simulation as <!--rinline meanVar -->, which is close to 
     the theoretical variance of 25.</p>

  <p>Distribution of 10000 exponentials vs distribution of 1000 means of 40 exponentials:</p>
  
```{r}
# Distribution of 10000 exponentials
hist(rexp(10000), col='salmon', border = 'salmon')
# Distribution of 1000 means of 40 exponentials with normal curve overlayed 
ggplot(data=expo.df, aes(x=allMeans)) + 
    geom_histogram(aes(y=..density..), binwidth =.1, fill='salmon', alpha=0.5) + 
    geom_vline(xintercept = mean, color='darkred',linetype='dashed') + 
    ggtitle('Distribution of 1000 simulation of means \nof 40 r.v. under exponential distribution') + 
    theme(plot.title = element_text(size = 10)) + theme_minimal() + 
    stat_function(fun=dnorm, args= list(mean=apply(expo.df,2,mean), sd=apply(expo.df, 2, sd)))
```

  <p>The distribution of 1000 means of 40 exponentials (2nd graph) appears to be normally distributed, whereas the distribution of 10000 exponentials (1st graph) does not.</p>
  </body></html>
