---
title: "STAT4007 QUALITY CONTROL HW3"
date: "4/9/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Overview: This assignment is to construct tabular CUSUM, CUSUM chart, tabular EWMA 
and EWMA chart for a chemical process, which has bath concentration measurements 
in ppm for the last 30 hours. The target mean value mu naught is 175ppm.

Load and prepare the data 
```{r}
mydata <- read.table("~/Downloads/hw3prob5data.txt")
myvector <- as.vector(mydata$V1)
# estimate sigma from the standard deviation of 30 data points
sigma <- sd(myvector) 
# standardized parameters h and k 
h <- 5
k <- 0.5
mu0 <- 175
head(myvector)
```
```{r}
K <- k*sigma
H <- h*sigma
n <- length(myvector)
Cu <- myvector
Cl <- myvector
Nu <- rep(0,n)
Nl <- rep(0,n)
Cu[1] <- max(0, myvector[1]- (mu0+K))
Cl[1] <- max(0, (mu0-K)-myvector[1])
Nu[1] <- (Cu[1]>0)
Nl[1] <- (Cl[1]>0)
for(i in 2:n){
    Cu[i] <- max(0, myvector[i]-(mu0+K)+Cu[i-1])
    Cl[i] <- max(0, (mu0-K)-myvector[i]+Cl[i-1])
    Nu[i] <- ((Nu[i-1]+1)*(Cu[i]>0))
    Nl[i] <- ((Nl[i-1]+1)*(Cl[i]>0))
}
```
```{r}
plot(Cu, type = "n", ylab = "Cumulative Sum", xlab = NA, main = "Cusum Status Chart")
lines(Cu, lty=1)
lines(Cl, lty=2)
points(Cu,pch=4)
points(Cl,pch=6)
legend(10,200, c("Upper Cusum","Lower Cusum"),pch=c(4,6),cex=0.9)
abline(h=H,lty=6)
text(5,100, "decision interval H",cex=.8)
```

From the above chart, the process is out-of-control at the 26th sample.
Construct a cusum status table 
```{r}
table1 <- cbind("i" = seq(1,n), "xi"=myvector, "xi-(mu0+K)"= myvector-(mu0+K), "Ci+" = Cu, "Ni+"=Nu, "(mu0-K)-xi"=(mu0-K)-myvector, "Ci-"=Cl, "Ni-"= Nl)
head(table1)
```

The same chart and table constructed with a 50% headstart:
```{r}
# HEADSTART
mydata <- read.table("~/Downloads/hw3prob5data.txt")
myvector <- as.vector(mydata$V1)
sigma <- sd(myvector)
h <- 5
k <- 0.5
mu0 <- 175
K <- k*sigma
H <- h*sigma
n <- length(myvector)
Cu <- myvector
Cl <- myvector
Nu <- rep(0,n)
Nl <- rep(0,n)
Cu[1] <- max(0, myvector[1]- (mu0+K)+H/2)
Cl[1] <- max(0, (mu0-K)-myvector[1]+H/2)
Nu[1] <- (Cu[1]>0)
Nl[1] <- (Cl[1]>0)
for(i in 2:n){
    Cu[i] <- max(0, myvector[i]-(mu0+K)+Cu[i-1])
    Cl[i] <- max(0, (mu0-K)-myvector[i]+Cl[i-1])
    Nu[i] <- ((Nu[i-1]+1)*(Cu[i]>0))
    Nl[i] <- ((Nl[i-1]+1)*(Cl[i]>0))
}
```
```{r}
plot(Cu, type = "n", ylab = "Cumulative Sum", xlab = NA, main = "Cusum Status Chart")
lines(Cu, lty=1)
lines(Cl, lty=2)
points(Cu,pch=4)
points(Cl,pch=6)
legend(10,200, c("Upper Cusum","Lower Cusum"),pch=c(4,6),cex=0.9)
abline(h=H,lty=6)
text(5,100, "decision interval H",cex=.8)
```
```{r}
table2 <- cbind("i" = seq(1,n), "xi"=myvector, "xi-(mu0+K)"= myvector-(mu0+K), "Ci+" = Cu, "Ni+"=Nu, "(mu0-K)-xi"=(mu0-K)-myvector, "Ci-"=Cl, "Ni-"= Nl)
```
Constructing Exponentially-Weighted Moving Average Chart using λ=0.05 and L=2.6:
```{r}
# Prepare varaibles and plug into formulas
lambda <- 0.05
L <- 2.6
z <- myvector
z[1] <- lambda*myvector[1]+(1-lambda)*mu0
for (i in 2:n){
    z[i] <- lambda*myvector[i]+(1-lambda)*z[i-1]
    UCL <- mu0 + L*sigma*sqrt((lambda/(2-lambda))*(1-(1-lambda)^(2*seq(1,n))))
    LCL <- mu0 - L*sigma*sqrt((lambda/(2-lambda))*(1-(1-lambda)^(2*seq(1,n))))
}
```

```{r}
# plot the graph
plot(z,type="b",pch=23,main="EWMA Chart",ylim=c(min(LCL,z),max(UCL,z)))
abline(h=mu0) 								
lines(UCL,type="l",pch=23)						
lines(LCL,type="l",pch=23)			
larger<-as.vector(z>UCL)
smaller<-as.vector(z<LCL)
```

```{r}
# construct EWMA in tabular form 
table3 <- cbind('i'=seq(1,n), "xi"= myvector, "EWMA, zi"=z, "UCL"=UCL, "LCL"=LCL, ">UCL?"=larger, "<LCL?"=smaller)
head(table3)
```